R-LIME: LIME法における矩形制約と最適化

### 2024-03-04 paper/paper.tex
Completed graduation thesis

% set the path for figures
\graphicspath{{./data/}{../python/data/}}

@@

% settings for cleveref
\crefname{figure}{Fig.}{Figs.}
\crefname{table}{Table}{Tables}
\crefname{algorithm}{Algorithm}{Algorithms}

@@

\section{Proposed Method}
\begin{figure}[tbp]
	\centering
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-lime}
		\caption{%
			LIME: Locally approximates the decision boundary around the focal point.
		}\label{fig:lime}
	\end{subfigure}%
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-anchor}
		\caption{%
			Anchor: Maximizes coverage of a rectangular region
			containing the focal point under accuracy constraints.
		}\label{fig:anchor}
	\end{subfigure}
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-rlime}
		\caption{%
			R-LIME:
			Maximizes coverage of a rectangular region containing the focal point
			under lower constraints on the accuracy of the linear classifier.
		}\label{fig:rlime}
	\end{subfigure}
	\caption[Visual comparison of LIME, Anchor and R-LIME]{%
		Visual comparison of LIME, Anchor and R-LIME (our proposed method).
		The solid line represents the rectangular region containing the focal
		point, and the dashed line represents the learned approximation model.
	}
\end{figure}
\subsection{Previous Work}
We specifically focuses on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.
\subsubsection[LIME]{%
	LIME (Local Interpretable Model-agnostic Explanations)
	\cite{ribeiro2016why}
}
LIME locally
approximates a trained black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The process involves:
\begin{enumerate}
	\item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$
	      and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
	\item Learning a linear classifier $g$
	      using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
	      by minimizing the following loss function:
	      \begin{equation}
		      \label{eq:lime_loss}
		      \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
		      \pi_x(z){\left(f(z)-g(z)\right)}^2,
	      \end{equation}
	      where $\pi_x(z)$ is a weight function designed to be larger for samples
	      closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}

@@

\subsection{Algorithm}\label{sec:alg}
{%
	\begin{figure}[t]
		\centering
		\includegraphics[width=0.32\textwidth]{rlime1}
		\includegraphics[width=0.32\textwidth]{rlime2}
		\includegraphics[width=0.32\textwidth]{rlime3}
		\caption[Overview of the R-LIME algorithm]{%
			Overview of the R-LIME algorithm.
			The progression of the algorithm is illustrated from left to right.
			The solid line represents the rectangular region $A$,
			and the dashed line represents the linear approximation model $g$
			learned within $A$.
			The initial value of $A$ is an empty rule (entire input space),
			and predicates are added to $A$, reducing coverage.
			The process continues until $\Prec(A)\ge\tau$,
			at which point the rule with the maximum coverage is output.
		}
	\end{figure}
	\def\myidt{\hspace{\algorithmicindent}}
	\ifnum\mode=0
		\begin{algorithm}[t]
			\small
			\input{alg1}
		\end{algorithm}
	\else
		\begin{algorithm}[p]
			\small
			\input{alg1}
		\end{algorithm}
	\fi

@@

\subsection{Computational Complexity}
Post-hoc explanation methods including LIME, Anchor, and R-LIME
need to query the black-box model multiple times
to get labels for perturbed samples, which is computationally expensive.
The number of queries required for LIME is $|\mathcal{Z}_p|$,
where $|\mathcal{Z}_p|$ is the number of perturbed samples.
On the other hand,
the number of queries required for Anchor and R-LIME is
$\mathcal{O}(B\cdot m^2+m^2\cdot\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]})$,
where $B$ is the beam width, $m$ is the number of features, and
$\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ is the computational complexity of
pure exploration
of the multi-armed bandit algorithm with $B\cdot m$ arms and beam width $B$.
$\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ depends on hyperparameters
such as threshold $\tau$, confidence level $1-\delta$ and tolerance $\epsilon$.

@@

Explanations generated by LIME (\cref{fig:A-lime,fig:B-lime}) indicate that
attributes such as having a significant prior record (Prior) or committing a
crime against property (Crime against Property) primarily contribute positively
to the prediction (predicting that the prisoner will be re-arrested).
On the other hand,
attributes like older age (Age), being married (Married),
and being of white race (Race) contribute negatively to the prediction
(predicting that the prisoner will not be re-arrested).
While these LIME explanations provide valuable insights into the behavior of
the black-box model,
they do not explicitly indicate the application scope of the explanations,
leaving users unable to determine to which inmates the explanations are applicable.

In contrast, R-LIME expresses the application scope of explanations
as a conjunction of predicates.
For example, the explanation for instance A
under $\tau=0.70$ (\cref{fig:A-rlime-70}) indicates that it is applicable
only to married prisoner (Married$=$Yes).
Furthermore, the generated explanations include the accuracy and coverage of
the approximation region, allowing users to evaluate how reliable the
explanations are.
For example, the coverage of the explanation for instance B under $\tau=0.90$
(\cref{fig:B-rlime-90}) is 0.002,
indicating that the decision boundaries around instance B are complex,
making it challenging to obtain a high-accuracy linear approximation.
This allows users to discern that the application scope of this explanation is
very narrow, limiting its utility.

@@

\section{Challenges and Future Perspectives}
 {%
  \def\imgwidth{0.45\textwidth}
  \begin{figure}[t]
	  \centering
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[width=0.7\textwidth]{newlime}
		  \caption{R-LIME for balanced label distribution.}\label{fig:balanced}
	  \end{subfigure}
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[width=0.7\textwidth]{newlime_for_imbalanced_data}
		  \caption{R-LIME for imbalanced label distribution.
		  }\label{fig:imbalanced}
	  \end{subfigure}
	  \caption[Behavior of R-LIME for balanced and imbalanced label distribution]{%
		  Behavior of R-LIME for balanced and imbalanced label distribution.
		  In case of imbalanced label distribution,
		  the approximation region covers the entire input space and the
		  linear approximation model always outputs the majority label.
	  }
  \end{figure}
 }

>>>>

% set the path for figures
\graphicspath{{./src/}{../python/data/}{../experiments/}}

@@

% settings for cleveref
\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\crefname{algorithm}{Algorithm}{Algorithms}

@@

\section{Proposed Method}
\begin{figure}[tbp]
	\centering
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-lime}
		\caption{%
			LIME: Locally approximates the decision boundary around the focal point.
		}\label{fig:lime}
	\end{subfigure}%
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-anchor}
		\caption{%
			Anchor: Maximizes coverage of a rectangular region
			containing the focal point under accuracy constraints.
		}\label{fig:anchor}
	\end{subfigure}
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-rlime3}
		\caption{%
			R-LIME:
			Maximizes coverage of a rectangular region
			under lower constraints on the accuracy of the linear classifier.
		}\label{fig:rlime}
	\end{subfigure}
	\caption[Visual comparison of LIME, Anchor and R-LIME]{%
		Visual comparison of LIME, Anchor and R-LIME (our proposed method).
		The solid line represents the rectangular region containing the focal
		point, and the dashed line represents the learned approximation model.
	}
\end{figure}
\subsection{Previous Work}
We specifically focuses on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.
\subsubsection[LIME]{%
	LIME (Local Interpretable Model-agnostic Explanations)
	\cite{ribeiro2016why}
}
LIME locally
approximates a trained black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The process involves:
\begin{enumerate}
	\item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$
	      and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
	\item Learning a linear classifier $g$
	      using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
	      by minimizing the following loss function:
	      \begin{equation}
		      \label{eq:lime_loss}
		      \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
		      \pi_x(z){\left(f(z)-g(z)\right)}^2,
	      \end{equation}
	      where $\pi_x(z)$ is a weight function designed to be larger for samples
	      closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}

@@

\subsection{Algorithm}\label{sec:alg}
{%
	\begin{figure}[t]
		\centering
		\includegraphics[width=0.32\textwidth]{visual-rlime1}
		\includegraphics[width=0.32\textwidth]{visual-rlime2}
		\includegraphics[width=0.32\textwidth]{visual-rlime3}
		\caption[Overview of the R-LIME algorithm]{%
			Overview of the R-LIME algorithm.
			The progression of the algorithm is illustrated from left to right.
			The solid line represents the rectangular region $A$,
			and the dashed line represents the linear approximation model $g$
			learned within $A$.
			The initial value of $A$ is an empty rule (entire input space),
			and predicates are added to $A$, reducing coverage.
			The process continues until $\Prec(A)\ge\tau$,
			at which point the rule with the maximum coverage is output.
		}
	\end{figure}
	\def\myidt{\hspace{\algorithmicindent}}
	\ifnum\mode=0
		\begin{algorithm}[t]
			\small
			\input{alg1}
		\end{algorithm}
	\else
		\begin{algorithm}[p]
			\small
			\input{alg1}
		\end{algorithm}
	\fi

@@

\ifnum\mode=1
	\subsection{Computational Complexity}
	Post-hoc explanation methods including LIME, Anchor, and R-LIME
	need to query the black-box model multiple times
	to get labels for perturbed samples, which is computationally expensive.
	The number of queries required for LIME is $|\mathcal{Z}_p|$,
	where $|\mathcal{Z}_p|$ is the number of perturbed samples.
	On the other hand,
	the number of queries required for Anchor and R-LIME is
	$\mathcal{O}(B\cdot m^2+m^2\cdot\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]})$,
	where $B$ is the beam width, $m$ is the number of features, and
	$\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ is the computational complexity of
	pure exploration
	of the multi-armed bandit algorithm with $B\cdot m$ arms and beam width $B$.
	$\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ depends on hyperparameters
	such as threshold $\tau$, confidence level $1-\delta$ and tolerance $\epsilon$.
\fi

@@

Explanations generated by LIME (\cref{fig:A-lime,fig:B-lime}) provide insights
that attributes such as having a prior offenses (Priors),
being served for a long time in prison (MonthsServed),
and committing a crime against property (Crime against Property)
primarily contribute to the positive prediction
(prediction that the prisoner will be re-arrested).
On the other hand,
attributes like older age (Age), being married (Married),
and being of white race (Race) contribute to the negative prediction
(prediction that the prisoner will not be re-arrested).
While these LIME explanations provide valuable insights into the behavior of
the black-box model,
they do not explicitly indicate the application scope of the explanations,
leaving users unable to determine to which inmates the explanations are applicable.

In contrast, R-LIME expresses the application scope of explanations
as a conjunction of predicates.
For example, the explanation for instance A
under $\tau=0.70$ (\cref{fig:A-rlime-70}) indicates that it is applicable
only to married prisoner (Married$=$Yes).
Furthermore, the generated explanations include the accuracy and coverage of
the approximation region, allowing users to evaluate how reliable the
explanations are.
For example, the coverage of the explanation for instance B under $\tau=0.90$
(\cref{fig:B-rlime-90}) is 0.01\%,
indicating that the decision boundaries around instance B are complex,
making it challenging to obtain a high-accuracy linear approximation.
This allows users to discern that the application scope of this explanation is
very narrow, limiting its utility.

@@

\section{Challenges and Future Perspectives}
 {%
  \def\imgwidth{0.45\textwidth}
  \begin{figure}[t]
	  \centering
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[width=0.8\textwidth]{visual-rlime3}
		  \caption{R-LIME for balanced label distribution.}\label{fig:balanced}
	  \end{subfigure}
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[width=0.8\textwidth]{visual-rlime-imbalanced}
		  \caption{R-LIME for imbalanced label distribution.
		  }\label{fig:imbalanced}
	  \end{subfigure}
	  \caption[Behavior of R-LIME for balanced and imbalanced label distribution]{%
		  Behavior of R-LIME for balanced and imbalanced label distribution.
		  In case of imbalanced label distribution,
		  the approximation region covers the entire input space and the
		  linear approximation model always outputs the majority label.
	  }
  \end{figure}
 }
